Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-04-08 talk/icooolps2009/paper.tex
switch armin and maciek to have alphabetical order. Try a new abstract (mostly by toon, again :-) ). It's not done but better than the old.

\numberofauthors{3}
\author{
\alignauthor Carl Friedrich Bolz\\
       \affaddr{Heinrich-Heine-Universität Düsseldorf}\\
       \affaddr{STUPS Group}\\
       \affaddr{Germany}\\
       \email{cfbolz@gmx.de}
\alignauthor Antonio Cuni\\
       \affaddr{University of Genova}\\
       \affaddr{DISI}\\
       \affaddr{Italy}\\
       \email{cuni@disi.unige.it}
\and
\alignauthor Armin Rigo\\
       \email{arigo@tunes.org}
\alignauthor Maciej Fijalkowski\\
       \email{fijal@merlinux.eu}
}
\maketitle

@@

\begin{abstract}
We present techniques for improving the results when a tracing JIT compiler is
applied to an interpreter.XXXAbrupt start - is the relevanz really immediately clear?XXX An unmodified tracing JIT performs not as well as one
would hope when the compiled program is itself a bytecode interpreter. We
examine the reasons for that, and how matters can be improved by adding markers to
the interpreter that help the tracing JIT to improve the results. We evaluate
the techniques by using them both on a small example as well as on a full Python
interpreter. This work has been done in the context of the PyPy project.

>>>>

\numberofauthors{3}
\author{
\alignauthor Carl Friedrich Bolz\\
       \affaddr{Heinrich-Heine-Universität Düsseldorf}\\
       \affaddr{STUPS Group}\\
       \affaddr{Germany}\\
       \email{cfbolz@gmx.de}
\alignauthor Antonio Cuni\\
       \affaddr{University of Genova}\\
       \affaddr{DISI}\\
       \affaddr{Italy}\\
       \email{cuni@disi.unige.it}
\and
\alignauthor Maciej Fijalkowski\\
       \email{fijal@merlinux.eu}
\alignauthor Armin Rigo\\
       \email{arigo@tunes.org}
}
\maketitle

@@

\begin{abstract}
XXX the first paragraph is a bit too long, the second too strange, but I think
it is a step in the right direction.

Tracing JIT compilers can greatly speed up programs that spend most of their time
in loops in which they take similar code paths.  Applying an unmodified tracing
JIT to an application that is itself a bytecode interpreter results in very
limited or no speedup. While bytecode interpreters spend most of their time in a
loop, \ie the bytecode dispatch, the code paths taken depend fully on the input
bytecode which the interpreter evaluates and thus changes all the time.

In this paper we show how to guide tracing JIT compilers to greatly improve
the speed of bytecode interpreters towards input bytecode. We evaluate our
technique by applying it to both a small example as well as to a full Python 
interpreter. This research has been conducted in the context of the PyPy 
project.
